### 2021-04-03 Chapter-3/Chapter-3.tex
update
Updates from Overleaf

Figure~\ref{fig:chapter3:cf}(b) shows a different situation. While the sensitive input \textsf{k} can affect the if-else branch, the compiler removes the branch with a conditional set instruction.  The opposite situation is the same. For example, recent work tries to use bit masking to rewrite the program with control flows. However, for some situations~\cite{Coppens:2009:PMT:1607723.1608124}, compilers (e.g., GCC) optimize the code too much (from a security view) and remove the unnecessary copy by reintroducing conditional moves.
\subsubsection{Data Flow}
\begin{figure}[h]
\begin{minipage}{0.4\linewidth}
\begin{lstlisting}[xleftmargin=.0\textwidth, xrightmargin=.0\textwidth, frame=none]
uint8_t T[128];
...
int example3(uint8_t k, uint32_t m){
  uint32_t index = m;
  index = (index + k) % 128;
  uint8_t t = T[index];
  ...
}
\end{lstlisting}
\end{minipage}
\hfill
\begin{minipage}{0.4\linewidth}
\begin{lstlisting}[xleftmargin=.0\textwidth, xrightmargin=.00\textwidth, frame=none, numbers=none, mathescape=true]
...
lea    edx, [addr_T]
mov    eax, [addr_k]
add    eax, [addr_m]
and    eax, 0x7f
movzx  eax, [addr_T + eax*1]
...
\end{lstlisting}
\end{minipage}\caption*{(a) A True Leakage}

\begin{minipage}{0.4\linewidth}
\begin{lstlisting}[xleftmargin=.0\textwidth, xrightmargin=.0\textwidth, frame=none]
uint8_t T[32];
...
int example4(uint8_t k, uint32_t m){
  uint32_t index = m;
  index = (index + k) % 32;
  uint8_t t = T[index];
  ...
}
\end{lstlisting}
\end{minipage}
\hfill
\begin{minipage}{0.4\linewidth}
\begin{lstlisting}[xleftmargin=.0\textwidth, xrightmargin=.00\textwidth, frame=none, numbers=none, mathescape=true]
...
lea    edx, [addr_T]
mov    eax, [addr_k]
add    eax, [addr_m]
and    eax, 0x3f
movzx  eax, [addr_T + eax*1]
...
\end{lstlisting}
\end{minipage}\caption*{(b) A False Positive}
\caption{Secret-dependent Data Accesses}\label{fig:chapter3:da}
\end{figure}

@@

\section{Method}
\begin{figure}
    \centering
    \includegraphics[width=.3\columnwidth]{./figures/chapter3/attack.pdf}
    \caption{A side-channel attack}\label{fig:side-channel-attack}
\end{figure}

@@

If the memory access equals to $t_b$, we can use the constraint $c_i :
f_i(\beta, k, m) >> L = t_b >> L$ to model the observation on secret-dependent
data accesses. Let's take a look the examples in Figure~\ref{fig:chapter3:da}(a). 
Suppose the base address of the array \textsf{T}
is 10. We symbolize both the $k$ and $m$. The formula that represents the memory access at line 6 
is $(m + k) \mod 128$. Therefore, we have the formula: $(m + k_1) \mod 128 >> 6  \neq (m + k_2) \mod 128 >> 6$.
Here is one possible solution: $m = 0, k_1 = 0, k_2 = 127$. In fact, for any $m$, we can always find
possible $k_1$ and $k_2$ that satisfy the above inequation. It is a true leakage. The leakage can 
also be identified by previous tools like CacheD. Similarly, we can model the memory access at line 6
with the formula:  $(m + k_1) \mod 32 >> 6  \neq (m + k_2) \mod 32 >> 6$.

\section{Scalability}
\subsection{Trace-oriented Symbolic Execution}
While symbolic execution can capture fine-grained semantics of programs, it
is also notorious for its unbearable performance cost. Previous trace-oriented
symbolic execution based
works~\cite{203878,Chattopadhyay:2017:QIL:3127041.3127044} all have large
performance bottlenecks. As a result, those approaches either only apply to
small-size programs~\cite{Chattopadhyay:2017:QIL:3127041.3127044} or apply some
domain knowledge to simplify the analysis. 
%Those tools interpret each
%instruction and update memory cells and registers with formulas that
%captured the semantics of the execution and search different input values that
%can lead to different execution behaviors using constraint solver. 
We implement the approach presented in \S\ref{sec:trace-qif} and model the side-channels as formulas. While the tool can finish analyzing some simple cases like AES, it can
not handle complicated cases like RSA.
We observe that finding side-channels using symbolic execution is different from
traditional general symbolic execution and can be optimized to be as efficient
as other methods with approaches below.

\subsection{Interpret Instructions Symbolically}
Existing binary analysis tools~\cite{shoshitaishvili2016state,
10.1007/978-3-642-22110-1_37} usually translate machine instructions into
intermediate languages (IR) to simplify analysis. 
The reason is that the number of machine instructions is
enormous, and the semantics of each instruction is complex. Intel Developer
Manual~\cite{intelsys} introduces more than 1000 different x86 instructions. 
IR typically has fewer instructions compared to the original machine ISA\@.
However, the IR layer, which predigest the implementation
and reduce the workload of those tools, is not suitable for side-channels 
analysis. Memory-based side-channels are very low issues. So, in general,
IR-based or source code side-channels analyses are not accurate enough.
In many cases, compilers can use conditional moves or bitwise operations to eliminate
branches. Also, as some IRs are not a superset or a subset of ISA, 
it is hard to rule out conditional jumps introduced by IR and add real branches 
eliminated by IR transformations.

Moreover, the IR design introduces significant overhead~\cite{217563}.
Transferring machine instructions into IR is time-consuming. For example,
REIL IR~\cite{dullien2009reil}, adopted in CacheS~\cite{236338}, has multiple
transform processes, from binary to VEX IR, BAP IR, and finally REIL IR\@. 
Also, IR increases the total number of instructions. For example, x86
instruction \textit{test eax, eax} transfers into 18 REIL IR instructions. If we
assume the time of symbolically executing grows as the number of instructions increase, the
design of adopting IR layers can introduce large overhead.

\vspace*{2pt}
\textbf{Our Solution to Challenge C2:}
We adopt the approach from QSYM~\cite{217563} and implement the symbolic execution
directly on the top of x86 instructions. Table~\ref{scala:ir} shows that
eliminating the IR layer can reduce the number of instructions executed during
the analysis.

@@

Existing approach relies on satisfiability modulo theories (SMT) solvers (e.g,
Z3~\cite{DeMoura:2008:ZES:1792734.1792766}) to find satisfying $k_1$ and $k_2$.
We argue that while it is a universal approach to solving constraints with SMT
solvers, for constraints with the above formats, using custom heuristics and
testing is much more efficient in practice. Constraint solving is a decision
problem expressed in logic formulas. SMT solvers transfer the inputted SMT
formula into the boolean conjunctive normal form (CNF) and feed it into the
internal boolean satisfiability problem (SAT) solver. The translation process,
called ``bit blasting'', is time-consuming. Also, as the SAT problem is a
well-known NP-complete problem, it is also hard to deal when it comes to
realistic uses with huge formulas. Despite the rapid development of SMT solvers
in recent years, constraint solving remains one of the obstacles to achieve the
scalability for real-world crypto systems.

\vspace*{2pt}
\textbf{Our Solution to Challenge C3:}
Instead of feeding the formula $f_a(k_1) \neq f_a(k_2)$ into a SMT solver, we
just randomly pick up $k_1, k_2 \in K$ and test them if they can satisfy the
formula. Our solution is based on the following intuition. For most combination
of $(k_{1}, k_{2} )$, the formula $f_a(k_1) \neq f_a(k_2)$ holds. As long as
$f_a$ is not a constant function, such $k_1, k_2$ must exist. For example,
suppose each time we only have 5\% chance to find such $k_1, k_2$, then after we
test with different input combination with 100 times, we have $1 -
(1-0.05)^{100} = 99.6\%$ chance find such $k_1, k_2$. Such random algorithms
work well for our problem.

@@

In the section, we discuss the limitations, usages, and some future work. 
%
\tool{} works on native x86 execution traces. The design, which is very
precise in terms of true leakages compared to other static source code
method~\cite{197207,BacelarAlmeida:2013:FVS:2483313.2483334}, also has 
the common limitations of dynamic approaches. \tool{} may only cover part of the code.
  % Each time we only get one single execution trace. Therefore,
We may 
neglect some side-channel vulnerabilities not covered by the traces analyzed.
However, this is not a crucial problem for analyzing crypto libraries because crypto
libraries are designed to have the same code coverage for various inputs. Our
evaluation also confirms the above point. For symmetric encryptions, during our
evaluation there is no secret-dependent control-flow transfers. RSA
implementations have several secret-dependent control-flow transfers. 
  %But after we manually check those leakages cites. We find
Most of them are  %used for
bound checks, which do not leak much information and have negligible effects
on the whole code coverage as well.

>>>>

Figure~\ref{fig:chapter3:cf}(b) shows a different situation. While the sensitive input \textsf{k} can affect the if-else branch, the compiler removes the branch with a conditional set instruction.  The opposite situation is the same. For example, recent work tries to use bit masking to rewrite the program with control flows. However, for some situations~\cite{Coppens:2009:PMT:1607723.1608124}, compilers (e.g., GCC) optimize the code too much (from a security view) and remove the unnecessary copy by reintroducing conditional moves.
\subsubsection{Data Flow}
\begin{figure}[h]
\begin{minipage}{0.4\linewidth}
\begin{lstlisting}[xleftmargin=.0\textwidth, xrightmargin=.0\textwidth, frame=none]
uint8_t T[128];
...
int example3(uint8_t k, uint32_t m){
  uint32_t index = m;
  index = index + k % 128;
  uint8_t t = T[index];
  ...
}
\end{lstlisting}
\end{minipage}
\hfill
\begin{minipage}{0.4\linewidth}
\begin{lstlisting}[xleftmargin=.0\textwidth, xrightmargin=.00\textwidth, frame=none, numbers=none, mathescape=true]
...
lea    edx, [addr_T]
mov    eax, [addr_k]
and    eax, 0x7f
add    eax, [addr_m]
movzx  eax, [addr_T + eax*1]
...
\end{lstlisting}
\end{minipage}\caption*{(a) A True Leakage}

\begin{minipage}{0.4\linewidth}
\begin{lstlisting}[xleftmargin=.0\textwidth, xrightmargin=.0\textwidth, frame=none]
uint8_t T[32];
...
int example4(uint8_t k, uint32_t m){
  uint32_t index = m;
  index = index + k % 32;
  uint8_t t = T[index];
  ...
}
\end{lstlisting}
\end{minipage}
\hfill
\begin{minipage}{0.4\linewidth}
\begin{lstlisting}[xleftmargin=.0\textwidth, xrightmargin=.00\textwidth, frame=none, numbers=none, mathescape=true]
...
lea    edx, [addr_T]
mov    eax, [addr_k]
and    eax, 0x3f
add    eax, [addr_m]
movzx  eax, [addr_T + eax*1]
...
\end{lstlisting}
\end{minipage}\caption*{(b) A False Positive}
\caption{Secret-dependent Data Accesses}\label{fig:chapter3:da}
\end{figure}

@@

\section{Method}\label{chapter3:method}
\begin{figure}
    \centering
    \includegraphics[width=.3\columnwidth]{./figures/chapter3/attack.pdf}
    \caption{A side-channel attack}\label{fig:side-channel-attack}
\end{figure}

@@

If the memory access equals to $t_b$, we can use the constraint $c_i :
f_i(\beta, k, m) >> L = t_b >> L$ to model the observation on secret-dependent
data accesses. Let's take a look the examples in Figure~\ref{fig:chapter3:da}(a). 
Suppose the base address of the array \textsf{T} is 10. We symbolize both the $k$ and $m$. The formula that represents the memory access at line 6 is $10 + m + (k \mod 128)$. Therefore, we have the formula: $(10 + m + (k_1 \mod 128)) >> 6  \neq (10 + m + (k_2 \mod 128)) >> 6$.
Here is one possible solution: $m = 0, k_1 = 0, k_2 = 127$. In fact, for any $m$, we can always find possible $k_1$ and $k_2$ that satisfy the above inequation. It is a true leakage. The leakage can also be identified by previous tools like CacheD. Similarly, we can model the memory access at line 6 with the formula:  $10 + m + (k_1 \mod 32) >> 6  \neq 10 + m + (k_2 \mod 32) >> 6$. If $m = 0$, we can not find satisfiable $k_1$ and $k_2$ for the above inequation, which indicates that it will always access the same cache line at line 6. It is the conclusion from Cache. However, it is possible that for some $m \in M$, we can find such a pair of $k_1$ and $k_2$ ($m = 53, k_1 = 1, k_2 = 0$). If the actual public input is $0$, existing tools (e.g., CacheD) can miss such a vulnerability. On the other hand, our tool can identify the leakages precisely.

\section{Scalability}
\subsection{Trace-oriented Symbolic Execution}
Symbolic execution is notorious for its high performance cost. 
Previous trace-oriented symbolic execution
work~\cite{203878,Chattopadhyay:2017:QIL:3127041.3127044} has serious
performance bottlenecks. As a result, these approaches either apply only to
small programs~\cite{Chattopadhyay:2017:QIL:3127041.3127044} or require
domain knowledge~\cite{Wang:2007:NCD:1250662.1250723} to simplify the analysis. 
Those tools interpret each
instruction and update memory cells and registers with formulas that
captured the semantics of the execution and search different input values that
can lead to different execution behaviors using constraint solver. 
We implement the approach presented in \S\ref{chapter3:method} and model the side-channels as formulas. While the tools can analyze some simple cases such as AES, it cannot handle complicated examples such as RSA.
We observe that finding side-channels using symbolic execution differs from
traditional symbolic execution, and it can be optimized to be as efficient
as other methods.

\subsection{Interpret Instructions Symbolically}
Existing binary analysis \cready{tools}{frameworks}~\cite{shoshitaishvili2016state,
10.1007/978-3-642-22110-1_37, song2008bitblaze} translate machine instructions into
intermediate languages (IR) to simplify analysis since
the variety of machine instructions is
enormous, and their semantics is complex. The Intel Developer
Manual~\cite{intelsys} documents more than 1000 different x86 instructions. 
Unfortunately, the IR layer, which
reduces the workload of these tools, is not suitable for side-channels 
analysis because
IR-based or source code side-channels analyses do not represent the executed instructions accurate enough to analyze fully their control and memory accesses.
For example, a compiler may use conditional moves or bitwise operations to eliminate
branches. Also, as some IRs are not a superset or a subset of ISA, 
it is hard to rule out conditional jumps introduced by IR and add real branches 
eliminated by IR transformations.

Moreover, the IR causes significant overhead~\cite{217563}.
Translating machine instructions into IR is time-consuming. For example,
REIL IR~\cite{dullien2009reil}, adopted in CacheS~\cite{236338}, has multiple
transform processes, from binary to VEX IR, BAP IR, and finally REIL IR\@. 
Also, IR increases the total number of instructions. For example, x86
instruction \textit{test eax, eax} transfers into 18 REIL IR instructions.

\textbf{Our Solution:}
We abandoned IR and expended the effort to implement 
symbolic execution directly on x86 instructions. 
Table~\ref{scala:ir} shows that eliminating the IR reduces the number 
of instructions examined during analysis. Previous works~\cite{217563} also 
adopted a similar approach to speed up fuzzing. Our implementation differs
from that work in two aspects: 1) We use complete constraints. 2) We run the
symbolic execution on one execution path each time. Our approach is approximately 30 times faster than using an IR (transferring ISA into IR and
symbolically executing it).

@@

Existing approaches rely on satisfiability modulo theories (SMT) solvers 
(e.g., Z3~\cite{DeMoura:2008:ZES:1792734.1792766}) to find satisfying assignments to
$k_1$ and $k_2$.
While this is a universal approach to solving constraints, 
for constraints of this form, using custom 
heuristics and testing is much more efficient in practice. Constraint 
solving is a decision problem expressed in logic formulas. SMT solvers 
transfer the SMT formula into the boolean conjunctive normal 
form (CNF) and feed it into the internal boolean satisfiability 
problem (SAT) solver. The translation process, called ``bit blasting'', 
is time-consuming. Also, as the SAT problem is a well-known NP-complete 
problem, it is hard to deal when it comes to
practical uses with huge formulas. Despite the rapid improvement 
in SMT solvers in recent years, constraint solving remains one of 
the obstacles to scaling the analysis of real-world crypto-systems.

\textbf{Our Solution:}
Instead of feeding the formula $f_a(k_1) \neq f_a(k_2)$ into a SMT solver, we
randomly pick $k_1, k_2 \in K$ and test them if they satisfy the
formula. Our solution is based on the following intuition. For most combination
of $(k_{1}, k_{2} )$, $f_a(k_1) \neq f_a(k_2)$. As long as
$f_a$ is not a constant function, such $k_1, k_2$ must exist. For example,
suppose each time we only have 5\% chance to find such $k_1, k_2$, then after we
test with different input combination with 100 times, we have $1 -
(1-0.05)^{100} = 99.6\%$ chance find such $k_1, k_2$. This type of random algorithm
works well for our problem.

@@

In the section, we discuss the limitations, usages, and some future work. 
%
\tool{} works on native x86 execution traces. The design, which is very
precise in terms of true leakages compared to other static source code
method~\cite{197207,BacelarAlmeida:2013:FVS:2483313.2483334}, also has 
the common limitations of dynamic approaches. \tool{} may only cover part of the code.
  % Each time we only get one single execution trace. Therefore,
We may 
neglect some side-channel vulnerabilities not covered by the traces analyzed.
However, this is not a crucial problem for analyzing crypto libraries because crypto
libraries are designed to have the same code coverage for various inputs. Our
evaluation also confirms the above point. For symmetric encryption, during our
evaluation there is no secret-dependent control-flow transfers. RSA
implementations have several secret-dependent control-flow transfers. 
  %But after we manually check those leakages cites. We find
Most of them are  %used for
bound checks, which do not leak much information and have negligible effects
on the whole code coverage as well.
